Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-06-06 SS3_Krill_481.Rmd
add last changes in html

---
title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
author:
  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
  - César Cárdenas^[Instituto Antártico Chileno, Chile]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a stock assessment model. Being able to identify index data relative, systematize, integrate and estimate krill population variables in a context of change and fisheries management is vital for decision making. On the other hand, spatial consideration and ecosystem aspect are another vital component for the application of this kind of approach. To do this, we use Stock Synthesis (SS3) as a statistical platform, a modeling tool widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. This approach has three main objectives. Firstly, integrate fishing, environmental and ecological variables in an integrated model. On the other hand, consider the spatial heterogeneity of the krill population structure. Finally, evaluate the impact of biological and population assumptions on the performance of this type of integrated stock assessment models. The work approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions, which in turn, allows the adequate configuration of control rules that may be useful in the fisheries management applied by CCAMLR of this important resource in the Southern Ocean. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013]. This regional organization has tried to conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys. However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. There is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables in krill. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others. In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery biological and ecosystem history on krill population over the past 20 years, as well as environmental variables, and predator abundance.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.  Each stratum may have distinct characteristics that affect krill biology (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows you to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others. Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques. All this data was proceses in @Dornam2021 and length compositions (Suppl. Mat. 3); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics on krill in the WAP area.  To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell) (Suppl. Mat 3). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series (Figure \@ref(fig:peng)), we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment. Al this data and temporal scale is showed in Figure \@ref(fig:datas2).

### Environmental variable in modelling

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics. SS3 addresses these approaches. In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.

@@

### Predator components in modelling

@@

A conceptual model for krill in the Antarctic Peninsula includes five spatial strata. Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)). Ecological and environmental variables include krill biomass, predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The analysis of results and outputs uses R tools (@R-base) and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].

## Population dynamics Model

In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time. This model incorporates key biological parameters such as growth rates, mortality rates, recruitment, and spawning biomass. The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.

A typical state-space model for krill population dynamics can be represented as:

\[
N_t = N_{t-1} \cdot e^{(r - M)} + R
\]

Where:

\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population. This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment. More detail in main equation and model configuration can be found in @methot2020stock. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).

@@

To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding this configurations. In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate 48.1                |
|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s2    | "s1" without S-R relation                                                            |
|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
|    s6    | "s1" Ricker S-R relation estimated                                                   |
|    s7    | "s1" w/ set of parameters estimated in @SAM-2024/32                                   |

@@

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @WP2024/34

These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment. All this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

@@

All models tested was the final gradient of the  model was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive definite which shows that the performance is within what was expected for this type of analysis.

Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there is variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall).

Overall, the model's performance is robust, but there is room for improvement in capturing the observed variability in certain fleets and years. In residuals components for `s1`, `s01` and `s7`, the model collects the modes and proportions for each source of information that contains lengths, however, it is the fleet of predators that fails to replicate the data of juvenile individuals (Figure \@ref(fig:fitcom)).

- 2. Residual (visual and metrics)

- 3. Retrospective analysis (Mhon Parameter)

- 4. Likelihood profile

Main variables and forecasting can be found in Table \@ref(tab:mainvar)

@@

Los análisis retrospectivo, dan cuenta de diferencias de estimación (sub - sobreestimación) en los patrones entre modelos evaluados.

## Hindcast Cross-Validation and prediction skill

Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated by `r4ss:SS_doRetro()`. As a robust measure of prediction skill, we implemented the mean absolute scaled error (MASE). In brief, the MASE score scales the mean absolute. Regarding (A MASE score > 1 indicates that the average model forecasts are worse than a random walk. Conversely, a MASE score of 0.5 indicates that the model forecasts twice as
accurately as a naïve baseline prediction; thus, the model has prediction skill.

models S01, S1, S2, and S5 exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment. Models S3 and S4 show minimal to no significant autocorrelations, suggesting more random recruitment patterns. Model S6 shows moderate autocorrelations, indicating some predictability based on recent past values. These findings help understand the temporal dynamics of krill recruitment and inform future modeling efforts. (fig autocorrelation)
## Kobe (status)

@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern.

@@

The R0 likelihood component profiles for the SMA and HAKE models (Fig. 6) were developed using the function r4ss:SS_profile() (see https://github.com/jabbamodel/ss3diags for example R code). For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources. The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro). The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.

## Cross correlation

## Comparision outputs betwwen scenarios

Regarding to autocorrelation in krill recruitment in time series analized by scenarios, **S01** most ACF values fall within the significance limits, suggesting no significant autocorrelation, which indicates the series might be random or lack temporal dependency. In **S1**, several ACF values at the initial lags are significant, indicating the presence of autocorrelation in the early periods and suggesting temporal dependency. Similarly, **S2** shows significant autocorrelation in the initial lags, pointing to early temporal dependency as well. **S3** mostly stays within the significance bounds, with some exceptions at the initial lags, hinting at a slight autocorrelation in the early periods, but it is not very pronounced. **S4** mirrors S01, with ACF values within the significance limits, indicating no significant autocorrelation and suggesting randomness in the series. **S5** displays some significant autocorrelation in the initial lags, although less pronounced than in S1 and S2, suggesting a mild early temporal dependency. **S6** follows the pattern of S01 and S4, with no significant autocorrelation, implying a potentially random series. Lastly, **S7** shows ACF values mostly within the significance bounds, with a few exceptions at the initial lags, indicating a slight autocorrelation in the early periods, though not strongly pronounced.In summary, S01, S4, and S6 exhibit no significant autocorrelation, suggesting randomness.

@@

# DISCUSION

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior. By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013]. Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.

Beyond integrating available information for assessing krill population dynamics, we tested several scenarios to identify uncertainties associated with estimating key population variables that aimed to diagnose, evaluate, assess performance, and compare the tested models. This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.

Development of an integrated stock assessment like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021]. Integrated models are not new to krill in SO. For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments. In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modelling strategies currently in use to manage krill"* and he stated that the an integrated modelling framework *"has the potential to substantially improve the scientific basis for future krill management"*.

@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1. @Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula. Although these models had convergence issues, they represented significant advances in this type of framework. Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings. Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO).

>>>>

---
title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
author:
  - Mardones Mauricio^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]

date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a integrated stock assessment model. Being able to identify that sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management nd decision making. On the other hand, spatial consideration and ecosystem aspect are vital components to model dynamics population on krill. This approach has three main objectives. Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance. On the other hand, consider the spatial heterogeneity of the krill population structure. Finally, evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models. Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.

@@

Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities. On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a]. Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities. As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013]. This regional organization has tried to conducts different modelling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016]. Various stock assessments approach on krill have been conducted to provide management advice to CCAMLR using model-based methodologies. Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity. Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013]. However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions. There is a growing consensus on the need for use stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill. Integrating the complexity of krill population dynamics presents several challenges. Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension. Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics. Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity. Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others. In response to this need, we propose a conceptual model to conducted a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.

@@

The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021 (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows you to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations. This include data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was process in @Dornam2021; and length compositions from AMLR survey (provide by Ecosystem Science Division, NOAA) was handling according to the evaluation template requirements (Suppl. Mat. 2); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics on krill in the WAP area.  To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST). The model included spatial component (cell) as random effect (formula: ~1 | cell) (Suppl. Mat 3). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020]. All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

A conceptual model for krill in the Antarctic Peninsula includes the five spatial strata described previously. Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)). Ecological and environmental variables that have impact in krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.  The analysis of results and outputs we use R tools (@R-base) and the graphical interface of the SS3 it was through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit). Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).

### Environmental variable modelling in `SS3`

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock]. In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.

@@

### Predator components modelling in `SS3`

@@

All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

@@

To consider the uncertainty associated with of krill nature dimension, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding this configurations. In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population, and for comparative propose, we consider `s1` as base model.

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate 48.1                |
|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
|    s2    | "s1" without S-R relation                                                            |
|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
|    s6    | "s1" Ricker S-R relation estimated                                                   |
|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par                                  |

@@

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @Mardones2024par (unpublished).

@@

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there is variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).

@@

Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better. @Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)). To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we use the joint-index residual. Overall, `s1` have a good performance when evaluating the lengths compositions (14.2%), however  RMSE show a low predictive power with respect to the indices (71.3%). A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).

@@

For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources. The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro). The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.

Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)

## Comparision outputs between scenarios

In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions. We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years[@Perry2020] (Figure \@(fig:recdev)).

Regarding to autocorrelation in krill recruitment in time series analized by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment. Models `s3` and `s4` show minimal to no significant autocorrelations, suggesting more random recruitment patterns. Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary, 
`s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness. 
(Figure \@(fig:acf)).

To evaluate consistency of the source of information used in each of the tested scenarios, we verify its influence on the estimate of R0 through likelihood components and, like scenario 1, we verify that all scenarios show a high dependence on the composition lengths of lengths with respect to the estimate of the variable (as show the (Figure \@ref(fig:likeall)).

With respect to population variables such as spawning biomass, we identify differences in the different scenarios with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios. This is relevant given that these scenarios consider low spatial dependence, density-dependence spawning biomass with respect to recruitment, which is plausible for resources such as krill where environmental dynamics shape population dynamics (Figure \@ref(fig:longterm)).

We carried out a specific comparison between the most divergent scenarios that we tested in this case, `s1`,  `s01` and `s7` which according to the description are the aggregated spatial scenario, disaggregate spatial scenario and scenario with new set of parameters, respectively. All with ecosystem components such as predators and an environmental variable.In this sense, the impacts and difference on the estimates of recruitment,  R0, virginal biomass and relative biomass are high (Figure \@ref(fig:comparision)), demonstrating the impact of initial considerations and assumptions on population variables estimation.

@@

# DISCUSSION

The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. Development of an integrated stock assessment model like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021]. However, integrated models are not new to krill in SO. For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments. In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modelling strategies currently in use to manage krill"* and he stated that the an integrated modelling framework *"has the potential to substantially improve the scientific basis for future krill management"*.

@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1. @Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula. Although these models had convergence issues, they represented significant advances in this type of framework. Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings. Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO). In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.

Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior. By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution. This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
